Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube

### 2022-09-11 latex/Chapters/DRLSearch.tex
update

For this project, I have implemented the above algorithm in a class called DeepQLearner which I shall describe in details in section \ref{sec:codelearners}.

@@

Since my implementation of \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}). This leaves us with the question of how can we use the output of such a network to search for a solution. One simple answer (which I have tried, see e.g. subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$. The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
\\
\\
Another possibility is to use an algorithm which combines both the actions distribution and the value function to inform its search. The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm. In short, it generates at each step (and possibly in a distributed or multithreaded manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found). The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function, according to a trade-off which can be tuned via a hyper-parameter. In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions. Let me first present some pseudo-code for the paths generation, in particular the trade-off between actions-values (exploration-exploitation):
\\

@@

Then the actual \textbf{MCTS} algorithm is simply generating a new path from root to a new leaf at each iteration, until it finds the goal:

@@

I have implemented the exact \textbf{MCTS} procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} and outlined in the above pseudo code in a class called MonteCarloTreeSearch (see section \ref{sec:codesolvers} for details).

>>>>

\noindent For this project, I have implemented the above algorithm in a class called DeepQLearner which I shall describe in details in section \ref{sec:codelearners}.

@@

The \textbf{DQL} heuristics from \ref{alg:TheoryDQL} no longer only produce cost-to-go estimates as \textbf{DRL}, but rather a joint cost-to-go and distribution over actions (i.e. a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}). This leaves us with the question of how can we use the output of such a network to search for a solution. One simple answer (which I have tried, see e.g. subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$. The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
\\
\\
Another possibility is to use an algorithm combining the actions distribution and the value function to inform its search. The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm. It generates at each iteration (possibly in a distributed or multithreaded manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found). The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (allowing exploration) and value function (exploitation), according to a trade-off which can be tuned via a hyper-parameter $c$. To prevent following the same paths (which might lead the search far from the goal), the \textbf{MCTS} keeps track of actions already followed and downweight their probability so that subsequent paths explore different actions. Let me first present in pseudo-code the paths generation, in particular the trade-off between actions-values (exploration-exploitation):
\\

@@

\noindent The actual \textbf{MCTS} algorithm is simply generating a new path from root to a new leaf at each iteration, until the goal is found:

@@

\noindent I have implemented the exact \textbf{MCTS} procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} and outlined above (\ref{alg:TheoryMCTS}) in a class called MonteCarloTreeSearch (see section \ref{sec:codesolvers} for details).

### 2022-09-11 latex/Chapters/Implementation.tex
update

\begin{landscape}
\begin{figure}[H]
\centering
\includegraphics[scale=0.40]{./Figures/codebase}
%\decoRule
\caption[Codebase]{Code base}
\label{fig:Codebase}
\end{figure}
\end{landscape}

Let me now describe what each submodule does in more details:

\section{rubiks.core}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasecore}
%\decoRule
\caption[Codebase]{rubiks.core}
\label{fig:Codebasecore}
\end{figure}
This submodule contains base classes that make the code base easier to use, debug, and extend. It contains the following:
\begin{itemize}
\item \textbf{Loggable}: a wrapper around Python's logger which automatically picks up classes' names at init and format things (dict, series and dataframes in particular) in a nicer way.
\item \textbf{Parsable}: a wrapper around ArgumentParser, which allows to construct objects in the project from command line, to define dependencies between object's configurations and to help a bit with typing of configs. The end result is that you can pretty much pass **kw\_args everywhere and it just works.
\item \textbf{Factory}: a typical factory pattern. Concrete factories can just define what widget they produce and the factory will help construct them from **kw\_args (or command line, since Factory inherits from Parsable)
\end{itemize}

\section{rubiks.puzzle}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasepuzzle}
%\decoRule
\caption[Codebase]{rubiks.puzzle}
\label{fig:Codebasepuzzle}
\end{figure}
This submodule contains:
\begin{itemize}
\item \textbf{Puzzle}: a Factory of puzzles. It defines states and actions in the abstract, and provides useful functions to apply moves, shuffle, generate training sets, tell if a state is the goal, etc. Puzzle can manufacture the two following types of puzzles:
\item \textbf{SlidingPuzzle}. Implements the states and moves of the sliding puzzle.
\item \textbf{RubiksCube}. Implements the states and moves of the Rubik's cube.
\\
\\
In addition, this module contains a \textbf{Puzzled} base class which most classes below inherit from. That allow e.g. heuristics, search algorithms, solvers and learners to know what puzzle and dimension they operate on, without having to reimplement these basic facts in each of them.
\end{itemize}

\section{rubiks.search}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasesearch}
%\decoRule
\caption[Codebase]{rubiks.search}
\label{fig:Codebasesearch}
\end{figure}
This modules contains graph search strategies. I have actually reused the code I implemented for one of the AIPnT assignments here. It contains the following classes:
\begin{itemize}
\item \textbf{Node}: which contains the state of a graph, as well as link to the previous (parent) state, action that leads from the latter to the former and the cost of the path so far.
\item \textbf{SearchStrategy}, a Factory class which can instantiate the following three types of search strategies to find a path to a goal:
\item \textbf{BreadthFirstSearch}, which is obviously an optimal strategy, but not particularly efficient.
\item \textbf{DepthFirstSearch}, which is not an optimal strategy, and also generally not particularly efficient.
\item \textbf{AStar}, which is optimal, and as efficient as the heuristic it makes use of is.
\end{itemize}

\section{rubiks.heuristics}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebaseheuristics}
%\decoRule
\caption[Codebase]{rubiks.heuristics}
\label{fig:Codebaseheuristics}
\end{figure}
\label{HSS}
This module contains base class Heuristic, also a Factory. Heuristic can instantiate the following heuristics, which we can use in the AStar strategy from the previous section:
\begin{itemize}
\item \textbf{Manhattan}: This heuristic is specific to the SP. It simply adds for each tile (other than the empty tile) the $L_{1}$ distance between their current position and their target position. We can quickly see that this heuristic is admissible. Indeeed, each move displaces one (and one only) tile by one position. In other puzzles, it is generally the case that a move will affect the position of many tiles (e.g. RC). Therefore, if tiles could somehow freely move on top of one-another (that is, we remove the constraint that there can be at most one tile per compartment, the number of moves necessary to solve the \textbf{SP} would exactly be the Manhattan distance.
\\
\\
The reason why it is interesting to have a known admissible heuristic is that we can obvioulsy compare other heuristics (DL, DRL, DQL, etc) in terms of optimality.
\\
\\
I have also implemented an improvement to the Manhattan distance, which I shall call Manhattan++ in here (and in code logs and graphs) and can be activated by simply passing \textit{plus=True} to the Manhattan Heuristic (see example \ref{ASSS} as well as a thorougher performance comparison on the 2x5 \textbf{SP} in \ref{MHComp}). It is based on the concept of linear constraints that I read about in lecture notes from Washington University (\cite{SlidingPuzzleLectureNotes}). The idea is simply that when tiles are already placed on their target row or column, but not in the expected order, we will need to get the tiles out of the way of one another to get to the goal state, and this is not accounted for by the Manhattan distance. For instance, in the following 2x3 puzzle, the Manhattan distance does not account for the fact that, at best, tile \red 3 \black needs to get out of the way for tiles \blue 1 \black and \blue 2 \black to move, and then needs to get back in its row, adding a cost of 2 to the Manhattan distance.

@@

\section{rubiks.deeplearning}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasedeeplearning}
%\decoRule
\caption[Codebase]{rubiks.deeplearning}
\label{fig:Codebasedeeplearning}
\end{figure}
This module is a wrapper around Pytorch. It contains:
\begin{itemize}
\item \textbf{DeepLearning}: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seemlessly passed to the networks and trained on. 
\item \textbf{FullyConnected}: wrapper around a Pytorch fully connected network, with configurable hidden layers and size. There are some params as well to add drop out, and to indicate whether or not the inputs are one hot encoding (in which case the first layer is automatically adjusted in size, using information from the puzzle dimension). There is also importantly a \textit{joint\_policy} parameter to indicate whether we want the output to be 1 dimensional (e.g. for value function learning) or $a+1$-dimensional, where $a$ is the size of the actions space (for joint policy-value learning).
\item \textbf{Convolutional}: similar wrapper to FullyConnected, but with the ability to add some parallel convolutional layers to complement fully connected layers.
\end{itemize}

\section{rubiks.learners}
\label{sec:codelearners}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebaselearners}
%\decoRule
\caption[Codebase]{rubiks.learners}
\label{fig:Codebaselearners}
\end{figure}
\label{PLcode}
\label{DRLcode}
\label{DQLcode}
This module implements learners, which learn something from a puzzle, store what they learnt, and can display interesting things about what they learnt.

@@

\section{rubiks.solvers}
\label{sec:codesolvers}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasesolvers}
%\decoRule
\caption[Codebase]{rubiks.solvers}
\label{fig:Codebasesolvers}
\end{figure}
This module implements solvers, which solve puzzles. The base class Solver is a Factory of solvers, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (shuffling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (chosable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver). Ideally the benchmark solver to compute the optimality score should be one which is provably optimal, but in the case of the \textbf{RC} that was not possible so I used A$^{*}$ with KociembaHeuristic as my benchmark. For the \textbf{SP} I used the optimal A$^{*}$ with Manhattan++ heuristic.

\begin{itemize}
\item \textbf{DFSSolver} This solver is literally a few lines of code, extremely close to the pseudo code detailed in \ref{sec:TheoryBFSDFS}
\item \textbf{BFSSolver} Ditto, see \ref{sec:TheoryBFSDFS}
\item \textbf{AStarSolver} Ditto, see \ref{sec:TheoryAStar}
\item \textbf{NaiveSlidingSolver} The NaiveSolver is conceptually simple, but was far from trivial to implement and debug. Its functioning is detailed in the appendix section \ref{sec:Examples}. Basically, it does what a beginner player would do (certainly what I do when I play the sliding puzzle), solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
\item \textbf{KociembaSolver} This solver is a wrapper around two libraries. The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable which solves the 3x3x3 Rubik's. My  wrapper simply translates my cube representation in that accepted by these 2 solvers, and their result string into moves usable by my code. I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 implementation, as I have detailed in appendix \ref{sec:Kociemba}.
\item \textbf{MonteCarloSearchTreeSolver} TBI I want to implement this in August when working on the Rubiks' ... following the Rubik's paper fromMcAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470}
\end{itemize}

\section{rubiks.scripts}

\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasescripts}
%\decoRule
\caption[Codebase]{rubiks.scripts}
\label{fig:Codebasescripts}
\end{figure}

Finally it is worth noting that the code will save on disk a lot of data (e.g. the learners will save what they have learnt, e.g. a Pytorch network or a data base of optimal costs, the performance comparison will run solvers versus very many configurations of puzzles and save the results for later being able to display) etc... The base of the tree to save all this data can be chosen by setting up the "RUBIKSDATA" environment variable. If not, it will go somewhere in you HOME :)

>>>>

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.5]{./Figures/codebase}}
  \caption[Codebase]{Code base}
  \label{fig:Codebase}
\end{figure}

\noindent Let me now describe what each submodule does in more details:

\Section{rubiks.core}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebasecore}
%\decoRule
\caption[Codebase]{rubiks.core}
\label{fig:Codebasecore}
\end{figure}
This submodule contains base classes that make the code base easier to use, debug, and extend. It contains the following:
\begin{itemize}
\item \textbf{Loggable}: a wrapper around Python's logger to automatically picks up classes' names at init and format things nicely (dict, series and dataframes in particular).
\item \textbf{Parsable}: a wrapper around ArgumentParser, which allows to construct objects in the project from command line, to define dependencies between object's configurations and to help a bit with typing of configs. The end result is that you can pretty much pass **kw\_args everywhere and it just works.
\item \textbf{Factory}: a typical factory pattern. Concrete factories can just define what widget they produce and the factory will help construct them from **kw\_args (or command line, since Factory inherits from Parsable)
\end{itemize}

\Section{rubiks.puzzle}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebasepuzzle}
%\decoRule
\caption[Codebase]{rubiks.puzzle}
\label{fig:Codebasepuzzle}
\end{figure}
This submodule contains:
\begin{itemize}
\item \textbf{Puzzle}: a Factory of puzzles. It defines states and actions in the abstract, and provides useful functions to apply moves, shuffle, generate training sets, tell if a state is the goal, etc. Puzzle can manufacture the two following types of puzzles:
\item \textbf{SlidingPuzzle}. Implements the states and moves of the sliding puzzle.
\item \textbf{RubiksCube}. Implements the states and moves of the Rubik's cube.
\\
\\
In addition, this module contains a \textbf{Puzzled} base class which most classes below inherit from. That allow e.g. heuristics, search algorithms, solvers and learners to know what puzzle and dimension they operate on, without having to reimplement these basic facts in each of them.
\end{itemize}

\Section{rubiks.search}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebasesearch}
%\decoRule
\caption[Codebase]{rubiks.search}
\label{fig:Codebasesearch}
\end{figure}
This modules contains graph search strategies. I have actually reused the code I implemented for one of the AIPnT assignments here. It contains the following classes:
\begin{itemize}
\item \textbf{Node}: which contains the state of a graph, as well as link to the previous (parent) state, action that leads from the latter to the former and the cost of the path so far.
\item \textbf{SearchStrategy}, a Factory class which can instantiate the following three types of search strategies to find a path to a goal:
\item \textbf{BreadthFirstSearch}, which is obviously an optimal strategy, but not particularly efficient.
\item \textbf{DepthFirstSearch}, which is not an optimal strategy, and also generally not particularly efficient.
\item \textbf{AStar}, which is optimal, and as efficient as the heuristic it makes use of is.
\end{itemize}

\Section{rubiks.heuristics}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebaseheuristics}
%\decoRule
\caption[Codebase]{rubiks.heuristics}
\label{fig:Codebaseheuristics}
\end{figure}
\label{HSS}
This module contains base class Heuristic, also a Factory. Heuristic can instantiate the following heuristics, which we can use in the AStar strategy from the previous section:
\begin{itemize}
\item \textbf{Manhattan}: This heuristic is specific to the SP. It simply adds for each tile (other than the empty tile) the $L_{1}$ distance between their current position and their target position. We can quickly see that this heuristic is admissible. Indeeed, each move displaces one (and one only) tile by one position. In other puzzles, it is generally the case that a move will affect the position of many tiles (e.g. RC). Therefore, if tiles could somehow freely move on top of one-another (that is, we remove the constraint that there can be at most one tile per compartment, the number of moves necessary to solve the \textbf{SP} would exactly be the Manhattan distance.
\\
\\
The reason why it is interesting to have a known admissible heuristic is that we can obvioulsy compare other heuristics (DL, DRL, DQL, etc) in terms of optimality.
\\
\\
I have also implemented an improvement to the Manhattan distance, which I shall call Manhattan++ in here (and in code logs and graphs) and can be activated by simply passing \textit{plus=True} to the Manhattan Heuristic (see example \ref{ASSS} as well as a thorougher performance comparison on the 2x5 \textbf{SP} in \ref{MHComp}). It is based on the concept of linear constraints that I read about in lecture notes from Washington University (\cite{SlidingPuzzleLectureNotes}). The idea is simply that when tiles are already placed on their target row or column, but not in the expected order, we will need to get the tiles out of the way of one another to get to the goal state, and this is not accounted for by the Manhattan distance. For instance, in the following 2x3 puzzle, the Manhattan distance does not account for the fact that, at best, tile \red 3 \black needs to get out of the way for tiles \blue 1 \black and \blue 2 \black to move, and then needs to get back in its row, adding a cost of 2 to the Manhattan distance.

@@

\Section{rubiks.deeplearning}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebasedeeplearning}
%\decoRule
\caption[Codebase]{rubiks.deeplearning}
\label{fig:Codebasedeeplearning}
\end{figure}
This module is a wrapper around Pytorch. It contains:
\begin{itemize}
\item \textbf{DeepLearning}: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seemlessly passed to the networks and trained on. 
\item \textbf{FullyConnected}: wrapper around a Pytorch fully connected network, with configurable hidden layers and size. Parameters allow to add drop out, to indicate whether or not the inputs are one hot encoding (in which case the first layer is automatically adjusted in size, using information from the puzzle dimension). There is also importantly a \textit{joint\_policy} parameter to indicate whether we want the output to be 1 dimensional (e.g. for value function learning) or $a+1$-dimensional, where $a$ is the size of the actions space (for joint policy-value learning).
\item \textbf{Convolutional}: similar wrapper to FullyConnected, but with the ability to add some parallel convolutional layers to complement fully connected layers.
\end{itemize}

\Section{rubiks.learners}
\label{sec:codelearners}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebaselearners}
%\decoRule
\caption[Codebase]{rubiks.learners}
\label{fig:Codebaselearners}
\end{figure}
\label{PLcode}
\label{DRLcode}
\label{DQLcode}
This module implements learners, which learn something from a puzzle, store what they learnt, and can display interesting things about what they learnt.

@@

\Section{rubiks.solvers}
\label{sec:codesolvers}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebasesolvers}
%\decoRule
\caption[Codebase]{rubiks.solvers}
\label{fig:Codebasesolvers}
\end{figure}
This module implements the actual puzzles solvers. The base class Solver is a Factory, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (scrambling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (configurable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver). Ideally the benchmark solver to compute the optimality score should provably optimal, but in the case of the \textbf{RC} that was not possible so I used A$^{*}$ with KociembaHeuristic as my benchmark. For the \textbf{SP} I used the optimal A$^{*}$ with Manhattan++ heuristic.

\begin{itemize}
\item \textbf{DFSSolver} This solver is literally a few lines of code, extremely close to the pseudo code detailed in \ref{alg:TheoryBFSDFS}
\item \textbf{BFSSolver} Ditto, see \ref{alg:TheoryBFSDFS}
\item \textbf{AStarSolver} Ditto, see \ref{alg:TheoryAStar}
\item \textbf{NaiveSlidingSolver} The NaiveSolver is conceptually simple, but was far from trivial to implement and debug. Its functioning is detailed in the appendix section \ref{sec:Examples}. Basically, it does what a beginner player would do, certainly what I do when I play the sliding puzzle: solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
\item \textbf{KociembaSolver} This solver is a wrapper around two libraries. The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable library which solves the 3x3x3 Rubik's. My  wrapper simply translates my cube representation in that accepted by these 2 solvers, and their result string into moves usable by my code. I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 Python library, as I have detailed in appendix \ref{sec:Kociemba}.
\item \textbf{MonteCarloSearchTreeSolver} My implementation follows McAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470} but is however single-threaded. As a consequence, I dropped the $\nu$ parameter from their implementation, since it only introduces a penalty to discourage parallel threads from generating the same paths (in single threaded mode, this parameter is inconsequential).
\\
This solver basically generates \textit{pseudo-random} paths from initial node to a new leaf at each iteration (see \ref{alg:TheoryMCTS}), until we obtain a goal leaf.Paths are, despite the name, completely deterministic: they indeed result from a trade-off between the probability of actions (penalised for repeated use) and the value function, but there are no random draws being performed.
\\ As per McAleer \& Agostilenni's paper, I have implemented a final pruning via \textbf{BFS} of the tree resulting from the above \textit{Monte-Carlo} process.It is inded clear that the paths generated by this \textbf{MCTS} algorithm have little chance of being optimal since there is no correction mechanism to favour shortcuts over long loops. I have however made this \textbf{BFS} trimming optional via a \textit{trim\_tree} parameter to study its effect on solutions' optimality and run time (see results section \ref{sec:MTCSHyperParamsEffect}).
\\
I will study and report on the effect of the trade-off hyper-parameter \textit{c}, and confirm how it affects solutions' quality and run time in section \ref{sec:MTCSHyperParamsEffect}.

\Section{data}

The code generates a fair amount of data, in particular:
\begin{itemize}
\item Manhattan++ generates databases of linear constraint penalties (see \ref{HSS})
\item TrainingData: (see e.g. \ref{DLSS}) generates databases of puzzles and their associated costs for later training of a DeepLearner.
\item Learners: DeepLearner, DeepReinforcementLearner, DeepQLearner generate models as well as some convergence data useful to understand how the networks learn. Similarly PerfectLearner saves its results for later reuse (in e.g. PerfectHeuristic) or display.
\item Solver: Generates sequences of scrambled puzzles used to make the performance test (comparing different solvers) fair.
\item Solver.performance\_test: Generates a data frame containing the metrics discussed earlier in \ref{sec:codesolvers}, for each solver and each level of difficulty.
\end{itemize}
That data is saved under a tree following the schema below:
$$\$\{ "RUBIKSDATA"\}/data\_type/puzzle\_type/dimension/file\_name.pkl$$

### 2022-09-11 latex/Chapters/Puzzles.tex
update

It is rather easy to see why this is the case, let us discuss why: given a configuration c of the tiles, let us define the permutation p(c) of this configuration according to the following schema: we enumerate the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment. For instance, the following 15-puzzle c:

@@

%-----------------------------------
%	SUBSECTION 1.2
%-----------------------------------
\Subsection{Search Space \& Solvability}

@@

 It is quite clear from that picture above that if we can bring the 4x4x4 cube to such a configuration on all faces (which is not super hard), and we know how to solve a 3x3x3, we are essentially done. (well, almost, since the additional degrees of freedom coming from the higher dimension of the 4x4x4 mean that certain arrangements of the colors which are possible on the 4x4x4 are not on the 3x3x3. For instance we can never have the center red and white on opposite faces of a 3x3x3, but we easily can on a 4x4x4! Fortunately that is nothing that a couple of easily learnable algorithms cannot fix).

>>>>

\noindent It is rather easy to see why this is the case, let us discuss why: given a configuration c of the tiles, let us define the permutation p(c) of this configuration according to the following schema: we enumerate the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment. For instance, the following 15-puzzle c:

@@

%-----------------------------------
%	SUBSECTION 1.2
%-----------------------------------
\Subsection{Search Space \& Solvability}
\label{Theory:SPSSS}

@@

\noindent It is quite clear from that picture above that if we can bring the 4x4x4 cube to such a configuration on all faces (which is not super hard), and we know how to solve a 3x3x3, we are essentially done. (well, almost, since the additional degrees of freedom coming from the higher dimension of the 4x4x4 mean that certain arrangements of the colors which are possible on the 4x4x4 are not on the 3x3x3. For instance we can never have the center red and white on opposite faces of a 3x3x3, but we easily can on a 4x4x4! Fortunately that is nothing that a couple of easily learnable algorithms cannot fix).
